---
title: "MyHomework"
author: "John Murphy"
output: html_document
date: "2023-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load Excel File and Add Header Row (10 points)

### Load Library, Load Excel File, and Add Header Row
```{r}
library(readxl)
Typical_Employee_Survey_Data <- read_excel('Typical_Employee_Survey_Data.xlsx', sheet='Survey', col_names = c( "Age", "Gender","Satisfaction", "Characteristics", "Years","Promoted", "Empowerment", "Budgetary", "Proud", "Turndown", "Relations"))
```

### Show Header Names
```{r}
names(Typical_Employee_Survey_Data)
```

### Show Beginning of File
```{r}
head(Typical_Employee_Survey_Data)
```

### Show End of File
```{r}
tail(Typical_Employee_Survey_Data)
```

## 2. Print out the 5-point summary and boxplot (numerical columns) for each column (20 points)

### Age Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Age)
fivenum(Typical_Employee_Survey_Data$Age)
boxplot(Typical_Employee_Survey_Data$Age)
```

### Gender Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Gender)
fivenum(Typical_Employee_Survey_Data$Gender)
boxplot(Typical_Employee_Survey_Data$Gender)
```

### Satisfaction Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Satisfaction)
fivenum(Typical_Employee_Survey_Data$Satisfaction)
boxplot(Typical_Employee_Survey_Data$Satisfaction)
```

### Characteristics Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Characteristics)
fivenum(Typical_Employee_Survey_Data$Characteristics)
boxplot(Typical_Employee_Survey_Data$Characteristics)
```

### Empowerment Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Empowerment)
fivenum(Typical_Employee_Survey_Data$Empowerment)
boxplot(Typical_Employee_Survey_Data$Empowerment)
```

### Promoted Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Promoted)
fivenum(Typical_Employee_Survey_Data$Promoted)
boxplot(Typical_Employee_Survey_Data$Promoted)
```

### Budgetary Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Budgetary)
fivenum(Typical_Employee_Survey_Data$Budgetary)
boxplot(Typical_Employee_Survey_Data$Budgetary)
```

### Proud Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Proud)
fivenum(Typical_Employee_Survey_Data$Proud)
boxplot(Typical_Employee_Survey_Data$Proud)
```

### Turndown Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Turndown)
fivenum(Typical_Employee_Survey_Data$Turndown)
boxplot(Typical_Employee_Survey_Data$Turndown)
```

### Relations Column
#### Show Summary, 5-point Summary, and Boxplot
```{r}
summary(Typical_Employee_Survey_Data$Relations)
fivenum(Typical_Employee_Survey_Data$Relations)
boxplot(Typical_Employee_Survey_Data$Relations)
```

## 3. Convert the columns that need to be converted to Factors (10 points)

### Gender Column
```{r}
Typical_Employee_Survey_Data$Gender <- as.factor(Typical_Employee_Survey_Data$Gender)
class(Typical_Employee_Survey_Data$Gender)
```

### Satisfaction Column
```{r}
Typical_Employee_Survey_Data$Satisfaction <- as.factor(Typical_Employee_Survey_Data$Satisfaction)
class(Typical_Employee_Survey_Data$Satisfaction)
```

### Characteristics Column
```{r}
Typical_Employee_Survey_Data$Characteristics <- as.factor(Typical_Employee_Survey_Data$Characteristics)
class(Typical_Employee_Survey_Data$Characteristics)
```

### Empowerment Column
```{r}
Typical_Employee_Survey_Data$Empowerment <- as.factor(Typical_Employee_Survey_Data$Empowerment)
class(Typical_Employee_Survey_Data$Empowerment)
```

### Promoted Column
```{r}
Typical_Employee_Survey_Data$Promoted <- as.factor(Typical_Employee_Survey_Data$Promoted)
class(Typical_Employee_Survey_Data$Promoted)
```

### Budgetary Column
```{r}
Typical_Employee_Survey_Data$Budgetary <- as.factor(Typical_Employee_Survey_Data$Budgetary)
class(Typical_Employee_Survey_Data$Budgetary)
```

### Proud Column
```{r}
Typical_Employee_Survey_Data$Proud <- as.factor(Typical_Employee_Survey_Data$Proud)
class(Typical_Employee_Survey_Data$Proud)
```

### Turndown Column
```{r}
Typical_Employee_Survey_Data$Turndown <- as.factor(Typical_Employee_Survey_Data$Turndown)
class(Typical_Employee_Survey_Data$Turndown)
```

### Relations Column
```{r}
Typical_Employee_Survey_Data$Relations <- as.factor(Typical_Employee_Survey_Data$Relations)
class(Typical_Employee_Survey_Data$Relations)
```

### Print first lines of table (showing column types)
```{r}
head(Typical_Employee_Survey_Data)
```


## 4. Plot the right visual for each column (20 points)

### Age Histogram
```{r}
hist(Typical_Employee_Survey_Data$Age, main="Histogram of Age vs People", col="Blue", xlab="Age", ylab="Number of People")
```

### Gender Plots
```{r}
plot(Typical_Employee_Survey_Data$Gender, col="Blue", main="Plot of Gender vs People", xlab="Gender (1=male, 2=female)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Gender, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Gender vs Age", xlab="Gender (1=male, 2=female)", ylab="Age")
```

### Satisfaction Plots
```{r}
plot(Typical_Employee_Survey_Data$Satisfaction, col="Blue", main="Plot of People vs Satisfaction", xlab="Satisfaction (1=very satisfied, 4=very dissatisfied)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Satisfaction, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Satisfaction", xlab="Satisfaction (1=very satisfied, 4=very dissatisfied)", ylab="Age")
```

### Characteristics Plots
```{r}
plot(Typical_Employee_Survey_Data$Characteristics, col="Blue", main="Plot of People vs Characteristics", xlab="Characteristics", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Characteristics, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Characteristics", xlab="Characteristics", ylab="Age")
```

### Years Histogram
```{r}
hist(Typical_Employee_Survey_Data$Years, main="Histogram of People vs Years", col="Blue", xlab="Years", ylab="Number of People")
```

### Promoted Plots
```{r}
plot(Typical_Employee_Survey_Data$Promoted, col="Blue", main="Plot of People vs Promoted", xlab="Promoted (1=very likely, 5=very unlikely)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Promoted, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Promoted", xlab="Promoted (1=very likely, 5=very unlikely)", ylab="Age")
```

### Empowerment Plots
```{r}
plot(Typical_Employee_Survey_Data$Empowerment, col="Blue", main="Plot of People vs Empowerment", xlab="Empowerment (1=always, 4=never)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Empowerment, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Empowerment", xlab="Empowerment (1=always, 4=never)", ylab="Age")
```

### Budgetary Plots
```{r}
plot(Typical_Employee_Survey_Data$Budgetary, col="Blue", main="Plot of People vs Budgetary", xlab="Budgetary (1=yes, 4=no)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Budgetary, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Budgetary", xlab="Budgetary (1=yes, 4=no)", ylab="Age")
```

### Proud Plots
```{r}
plot(Typical_Employee_Survey_Data$Proud, col="Blue", main="Plot of People vs Proud", xlab="Proud (1=very proud, 4=not at all proud)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Proud, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Proud", xlab="Proud (1=very proud, 4=not at all proud)", ylab="Age")
```

### Turndown Plots
```{r}
plot(Typical_Employee_Survey_Data$Turndown, col="Blue", main="Plot of People vs Turndown", xlab="Turndown (1=very likely, 5=very unlikely)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Turndown, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Turndown", xlab="Turndown (1=very likely, 5=very unlikely)", ylab="Age")
```

### Relations Plots
```{r}
plot(Typical_Employee_Survey_Data$Relations, col="Blue", main="Plot of People vs Relations", xlab="Relations (1=very good, 5=very bad)", ylab="Number of People")
plot(Typical_Employee_Survey_Data$Relations, Typical_Employee_Survey_Data$Age, col="Red", main="Plot of Age vs Relations", xlab="Relations (1=very good, 5=very bad)", ylab="Age")
```

## 5. Plot scatter plots to show few relationships (20 points)

### Age vs People Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Age, main="Scatter Plot of Age vs People", col="Blue", xlab="People", ylab="Age")
```

### Gender vs Age Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Age, as.numeric(Typical_Employee_Survey_Data$Gender), main="Scatter Plot of Gender vs Age", col="Blue", xlab="Age", ylab="Gender (1=male, 2=female)")
```

### Satisfaction and Gender vs People Scatter Plot
```{r}
plot(as.numeric(Typical_Employee_Survey_Data$Satisfaction), main="Scatter Plot of Gender and Satisfaction vs People", col="Red", xlab="People", ylab="Gender (Blue), Satisfaction (Red)")
points(as.numeric(Typical_Employee_Survey_Data$Gender), col="Blue")
```

### Satisfaction and Gender vs Age Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Age, as.numeric(Typical_Employee_Survey_Data$Satisfaction), main="Scatter Plot of Gender and Satisfaction vs Age", col="Red", xlab="Age", ylab="Gender (Blue), Satisfaction (Red)")
points(Typical_Employee_Survey_Data$Age, as.numeric(Typical_Employee_Survey_Data$Gender), col="Blue")
```

### Proud and Gender vs People Scatter Plot
```{r}
plot(as.numeric(Typical_Employee_Survey_Data$Proud), main="Scatter Plot of Gender and Proud vs People", col="Red", xlab="People", ylab="Gender (Blue), Proud (Red)")
points(as.numeric(Typical_Employee_Survey_Data$Gender), col="Blue")
```

### Proud and Gender vs Age Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Age, as.numeric(Typical_Employee_Survey_Data$Proud), main="Scatter Plot of Gender and Proud vs Age", col="Red", xlab="Age", ylab="Gender (Blue), Proud (Red)")
points(Typical_Employee_Survey_Data$Age, as.numeric(Typical_Employee_Survey_Data$Gender), col="Blue")
```

### Years vs Age Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Age, Typical_Employee_Survey_Data$Years, main="Scatter Plot of Years vs Age", col="Purple", xlab="Age", ylab="Years of Service")
```

### Years vs People Scatter Plot
```{r}
plot(Typical_Employee_Survey_Data$Years, main="Scatter Plot of Years vs People", col="Red", xlab="People", ylab="Years of Service")
```
